---
title: 'Week 10 Practice #2: Neural Networks'
author: "Jen McGregor"
date: "12/1/2021"
output:
  prettydoc::html_pretty:
  theme: hpstr
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidymodels)
library(keras)
library(ISLR)
```

# Problem 1

**This exercise uses the book’s Carseats data set, in which we are interested in predicting Sales.**

```{r}
myData <- Carseats
```


**1. Fit a neural network to the entire dataset. Report the cross-validated metrics.**

```{r}
set.seed(1)
nn_recipe <- recipe(Sales~., 
                     data = myData) %>%
  step_dummy(ShelveLoc)%>%
  step_dummy(Urban)%>%
  step_dummy(US) %>%
  step_normalize(all_numeric())

set.seed(1)
nn_cvs <- vfold_cv(myData, v = 5)

nn_mod <- mlp(
  hidden_units = 8,
  penalty = .01,
  epochs = 100,
  activation = "linear"
) %>%
  set_engine("nnet") %>%
  set_mode("regression")

nn_wflow <- workflow() %>%
  add_recipe(nn_recipe) %>%
  add_model(nn_mod)

nn_fit_1 <- nn_wflow %>%
  fit_resamples(nn_cvs)

nn_fit_1 %>% collect_metrics()
```

**2. Now, tune your neural network according to hidden_units and penalty to identify the best neural network model. Report the cross-validated metrics. Remember to consider the size of your dataset when specifying your model(s).**

```{r}
nn_grid <- grid_regular(
  hidden_units(c(2, 12)),
  penalty(c(-5, 0)),
  levels = 3
)

nn_mod <- mlp(
  hidden_units = tune(),
  penalty = tune(),
  epochs = 100,
  activation = "linear"
) %>%
  set_engine("nnet") %>%
  set_mode("regression")

nn_wflow <- workflow() %>%
  add_recipe(nn_recipe) %>%
  add_model(nn_mod)

nn_grid_search <-
  tune_grid(
    nn_wflow,
    resamples = nn_cvs,
    grid = nn_grid
  )

tuning_metrics <- nn_grid_search %>% collect_metrics() %>%
  filter(.metric=="rmse") %>%
  slice_min(mean)
tuning_metrics
```

**3. Are more hidden units necessarily better?**

In this example, increasing the amount of hidden units decreases the value of RMSE. Since smaller RMSE values are preferred, higher amounts of hidden units may be preferred. 


**4. How do these results compare to your previous results using decision trees and random forests?**

These results are both lower than our previous results using decision trees and random forests. 

