---
title: "Week 9 Activity #2"
author: "Jen McGregor"
output:
  prettydoc::html_pretty:
  theme: hpstr
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
library(tidyverse)
library(tidymodels)
library(kknn)
library(glmnet)
library(discrim)
library(kernlab)
set.seed(1)
```

## Data Prep

```{r include=FALSE}
# Read data
fed <- read.csv("https://www.dropbox.com/s/9t8sxr1sg0monih/federalist.txt?dl=1")

# Data from papers with known authors
fed_known <- fed %>%
  select(-1) %>%
  filter(Author != "DIS")
```

## Try it!

#### Fit a KNN classifier using step_pca.  How does the accuracy compare to when you use all the original predictors?

The KNN classifier using step_pca has a **higher** accuracy rate than a standard KNN classifier.

#### A. KNN Classifier with step_pca
```{r}
knn_mod_tune <- nearest_neighbor(neighbors = tune()) %>%
  set_engine("kknn") %>%
  set_mode("classification")
```

```{r}
knn_rec1 <- recipe(Author~.,data=fed_known) %>%
  step_normalize(all_numeric()) %>%
  step_pca(all_numeric(),threshold = .60)

knn_wflow <- workflow() %>%
  add_recipe(knn_rec1) %>%
  add_model(knn_mod_tune)

set.seed(1)
fed_known_cv <- vfold_cv(fed_known, v = 5, strata=Author)

k_grid <- grid_regular(neighbors())

knn_grid_search <-
  tune_grid(
    knn_wflow,
    resamples = fed_known_cv,
    grid = k_grid
  )

knn_grid_search %>% 
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  slice_max(mean)
```
#### B. Standard KNN Classifier

```{r, not PCA}
knn_rec2 <- recipe(Author~.,data=fed_known) %>%
  step_normalize(all_numeric())

knn_wflow <- workflow() %>%
  add_recipe(knn_rec2) %>%
  add_model(knn_mod_tune)

set.seed(1)
fed_known_cv <- vfold_cv(fed_known, v = 5,strata=Author)

k_grid <- grid_regular(neighbors())

knn_grid_search <-
  tune_grid(
    knn_wflow,
    resamples = fed_known_cv,
    grid = k_grid
  )

knn_grid_search %>% 
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  slice_max(mean)
```


## Try it!

#### Fit a support vector classifier, tuning the **cost** parameter

```{r}
svm_rec <- recipe(Author~., data=fed_known) %>%
  step_normalize(all_numeric())

svm_mod1 <- svm_poly(cost = tune(), degree = 1) %>%
  set_mode("classification") %>%
  set_engine("kernlab")

svm_grid <- grid_regular(cost())
svm_grid

svm_wflow <- workflow()%>%
  add_recipe(svm_rec) %>%
  add_model(svm_mod1)

set.seed(1)
fed_known_cv <- vfold_cv(fed_known, v = 5,strata=Author)

svm_grid_search <-
  tune_grid(
    svm_wflow,
    resamples = fed_known_cv,
    grid = svm_grid
  )

svm_grid_search %>% 
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  slice_max(mean)
```


#### Fit a support vector **machine**, tuning the **cost** parameter AND the **degree** parameter.

```{r}
svm_rec <- recipe(Author~., data=fed_known) %>%
  step_normalize(all_numeric())

svm_mod2 <- svm_poly(cost = tune(), degree = tune()) %>%
  set_mode("classification") %>%
  set_engine("kernlab")

svm_grid <- grid_regular(cost(),degree())
svm_grid

svm_wflow <- workflow()%>%
  add_recipe(svm_rec) %>%
  add_model(svm_mod2)

set.seed(1)
fed_known_cv <- vfold_cv(fed_known, v = 5,strata=Author)

svm_grid_search <-
  tune_grid(
    svm_wflow,
    resamples = fed_known_cv,
    grid = svm_grid
  )

svm_grid_search %>% 
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  slice_max(mean)
```

